### 2017-04-05 public.tex

\chapter*{在学期间论文发表情况}

\addcontentsline{toc}{chapter}{在学期间论文发表情况}

\renewcommand{\labelenumi}{[\arabic{enumi}].}
\begin{enumerate}
\item Jianmin Ji, Hai Wan, \textbf{Peng Xiao}. On Elementary Loops and Proper Loops for Disjunctive Logic Programs [C].//
Proceedings of the 29th AAAI Conference on Artificial
Intelligence (AAAI-15). 2015.
\item Jianmin Ji, Hai Wan, \textbf{Peng Xiao}, Ziwei Huo, Zhanhao Xiao. Elementary Loops Revisited [C].//
      Proceedings of the 28th AAAI Conference on Artificial
      Intelligence (AAAI-14). 2014.
\end{enumerate}