### 2017-04-06 public.tex

\chapter*{在学期间论文发表情况}

\addcontentsline{toc}{chapter}{在学期间论文发表情况}

\renewcommand{\labelenumi}{[\arabic{enumi}].}
\begin{enumerate}
\item 	Jianfeng Du, Hai Wan, \textbf{Huaguan Ma}. Practical TBox Abduction Based on Justification Patterns [C].//
Proceedings of the 31th AAAI Conference on Artificial
Intelligence (AAAI-17). 2017.
\end{enumerate}